# Data transformation

```{r, include=FALSE}
library(lubridate)
library(tidyverse)
library(dplyr)
library(zoo)
library(gridExtra)
library(ggplot2)
library(calendR)
```

## Metadata

For the metadata dataset, we filter out the features that we will use for visualization from the total 149 columns, focusing on date, ticket price, holiday type, temperature and event information. The date data is transformed into "%m/%d/%Y" format, which help us address it.

```{r, include=FALSE}
#load in metadata dataset
meta <- read_csv("https://cdn.touringplans.com/datasets/metadata.csv")
```

```{r}
#metadata transformation
bigmeta <- meta %>% 
  select(DATE,WDW_TICKET_SEASON,DAYOFWEEK,DAYOFYEAR,WEEKOFYEAR,MONTHOFYEAR, YEAR, HOLIDAYPX, HOLIDAYM,HOLIDAY, WDWMEANTEMP, WDWeventN,MKeventN,EPeventN,HSeventN,AKeventN)%>%
  mutate(DATE=as.Date(DATE, format=c("%m/%d/%Y")))

head(bigmeta)
```
## Attraction specific data

For the attraction specific datasets, the data is collected randomly nearly every 10 minutes. This also arise the problem of providing too many data points with about 110 rows each day. We transform the data into the tidy data by different level (hour, day, month, year). 

Besides, we notice that the date in attraction specific datasets is not continuous, there is missing date from 03/14/2020 to 06/26/2020 and from 06/28/2020 to 07/07/2020, from 07/21/2020 to 10/30/2020 and from 01/02/2021 to 04/29/2021. The reasonfor missing is that Disney World was closed in those days because of the Covid-19 pandemic. To address this problem, we fill the DATE to be complete according to the nature of date. 

Finally, we combine the 13 attraction specific data into one dataset, which may provide the option for us to make comparation between different attractions on the same date, and we may also facet on different attractions.

```{r, include=FALSE}
#Function for loading data depending on different uses
#Specify "by" for different average computations
#Default="everyday", if by="none", return the whole dataset
cleanthis <- function(someurl, by="everyday" ){
  somedata <- read_csv(someurl, show_col_types = FALSE)
  df <- somedata %>% select(-SACTMIN) %>% mutate(date=as.Date(datetime)) %>% filter(SPOSTMIN>=0)
  if (by=="year"){
    df <- df %>% mutate(year=year(date))%>% group_by(year) %>% summarize(mean_year = mean(SPOSTMIN))
  }else if(by=="month"){
      df <- df %>% mutate(month=month(date))%>% group_by(month) %>% summarize(mean_month = mean(SPOSTMIN))
  }else if(by=="day"){
      df <- df %>% mutate(day=day(date))%>% group_by(day) %>% summarize(mean_day = mean(SPOSTMIN))
  }else if(by=="everyday"){
      df <- df %>% group_by(date) %>% summarize(mean_everyday = mean(SPOSTMIN))
  }else if(by=="none"){
    df <- df
  }else if(by=="hour"){
    df$hour <- format(as.POSIXct(df$datetime), format = "%H")
  }
  return(df)
}
```


```{r, include=FALSE}
#Load all attraction waiting time datasets from the website, compute everyday mean
soarin <- cleanthis("https://cdn.touringplans.com/datasets/soarin.csv") %>%
  rename(soarin=mean_everyday)
avatar <- cleanthis("https://cdn.touringplans.com/datasets/flight_of_passage.csv") %>%
  rename(avatar=mean_everyday)
dinosaur <- cleanthis("https://cdn.touringplans.com/datasets/dinosaur.csv") %>%
  rename(dinosaur=mean_everyday)
ee <- cleanthis("https://cdn.touringplans.com/datasets/expedition_everest.csv") %>%
  rename(ee=mean_everyday)
safari <- cleanthis("https://cdn.touringplans.com/datasets/kilimanjaro_safaris.csv")%>% 
  rename(safari=mean_everyday)
navi <- cleanthis("https://cdn.touringplans.com/datasets/navi_river.csv")%>%
  rename(navi=mean_everyday)
pirate <- cleanthis("https://cdn.touringplans.com/datasets/pirates_of_caribbean.csv")%>% rename(pirate=mean_everyday)
rnrc <- cleanthis("https://cdn.touringplans.com/datasets/rock_n_rollercoaster.csv")%>%
  rename(rnrc=mean_everyday)
sevendwarf <- cleanthis("https://cdn.touringplans.com/datasets/7_dwarfs_train.csv")%>% rename(sevendwarf=mean_everyday)
slinkydog <- cleanthis("https://cdn.touringplans.com/datasets/slinky_dog.csv")%>%
  rename(slinkydog=mean_everyday)
spaceship <- cleanthis("https://cdn.touringplans.com/datasets/spaceship_earth.csv")%>%
  rename(spaceship=mean_everyday)
splash <- cleanthis("https://cdn.touringplans.com/datasets/splash_mountain.csv")%>%
  rename(splash=mean_everyday)
toystory <- cleanthis("https://cdn.touringplans.com/datasets/toy_story_mania.csv")%>% 
  rename(toystory=mean_everyday)
<<<<<<< HEAD

#by hour
soarin.hour <- cleanthis("https://cdn.touringplans.com/datasets/7_dwarfs_train.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
avatar.hour <- cleanthis("https://cdn.touringplans.com/datasets/flight_of_passage.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
dinosaur.hour <- cleanthis("https://cdn.touringplans.com/datasets/dinosaur.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
ee.hour <- cleanthis("https://cdn.touringplans.com/datasets/expedition_everest.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
safari.hour <- cleanthis("https://cdn.touringplans.com/datasets/kilimanjaro_safaris.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
navi.hour <- cleanthis("https://cdn.touringplans.com/datasets/navi_river.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
pirate.hour <- cleanthis("https://cdn.touringplans.com/datasets/pirates_of_caribbean.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
rnrc.hour <- cleanthis("https://cdn.touringplans.com/datasets/rock_n_rollercoaster.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
sevendwarf.hour <- cleanthis("https://cdn.touringplans.com/datasets/7_dwarfs_train.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
slinkydog.hour <- cleanthis("https://cdn.touringplans.com/datasets/slinky_dog.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
spaceship.hour <- cleanthis("https://cdn.touringplans.com/datasets/spaceship_earth.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
splash.hour <- cleanthis("https://cdn.touringplans.com/datasets/splash_mountain.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
toystory.hour <- cleanthis("https://cdn.touringplans.com/datasets/toy_story_mania.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
                      

whichyear <- c(2019,2021)

soarin.hour <- soarin.hour %>% filter(year %in% whichyear ) %>% select(year, SPOSTMIN)%>% mutate(ride = "soarin") 

avatar.hour <- avatar.hour %>% filter(year %in% whichyear) %>% select(year, SPOSTMIN)%>% mutate(ride = "avatar") 
  
dinosaur.hour <- dinosaur.hour %>% filter(year %in% whichyear) %>% select(year, SPOSTMIN)%>% mutate(ride = "dinosaur") 
  
ee.hour <- ee.hour %>% filter(year %in% whichyear) %>% select(year, SPOSTMIN)%>% mutate(ride = "ee") 

safari.hour <- safari.hour %>% filter(year %in% whichyear) %>% select(year, SPOSTMIN)%>% mutate(ride = "safari") 

navi.hour <- navi.hour %>% filter(year %in% whichyear) %>% select(year, SPOSTMIN)%>% mutate(ride = "navi") 

pirate.hour <- pirate.hour %>% filter(year %in% whichyear) %>% select(year, SPOSTMIN)%>% mutate(ride = "pirate") 

rnrc.hour <- rnrc.hour %>% filter(year %in% whichyear) %>% select(year, SPOSTMIN)%>% mutate(ride = "rnrc") 

sevendwarf.hour <- sevendwarf.hour %>% filter(year %in% whichyear) %>% select(year, SPOSTMIN)%>% mutate(ride = "sevendwarf") 

slinkydog.hour <- slinkydog.hour %>% filter(year %in% whichyear) %>% select(year, SPOSTMIN)%>% mutate(ride = "slinkydog") 

spaceship.hour <- spaceship.hour %>% filter(year %in% whichyear) %>% select(year, SPOSTMIN)%>% mutate(ride = "spaceship") 

splash.hour <- splash.hour %>% filter(year %in% whichyear) %>% select(year, SPOSTMIN)%>% mutate(ride = "splash") 

toystory.hour<- toystory.hour %>% filter(year %in% whichyear) %>% select(year, SPOSTMIN)%>% mutate(ride = "toystory") 

  
#Multiple Boxplot of year 2019
#allinone <- rbind(soarin.hour, avatar.hour,dinosaur.hour, ee.hour, safari.hour ,navi.hour, pirate.hour, rnrc.hour, sevendwarf.hour, slinkydog.hour, spaceship.hour, splash.hour, toystory.hour)
allinone <- rbind(soarin.hour, avatar.hour,dinosaur.hour, pirate.hour, rnrc.hour, sevendwarf.hour, slinkydog.hour, spaceship.hour, splash.hour, toystory.hour)

allinone %>% 
  ggplot(aes(x=reorder(ride,-SPOSTMIN, median), y=SPOSTMIN))+
  geom_boxplot(varwidth = TRUE)+
  facet_wrap(~year)+
  ggtitle("Multiple boxplot of all rides")+
  xlab("Ride name")+
  ylab("Waiting Time")+
  theme_grey(14)+
  theme(axis.text.x=element_text(size=9))


```
```{r}
soarin.hour <- cleanthis("https://cdn.touringplans.com/datasets/7_dwarfs_train.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
avatar.hour <- cleanthis("https://cdn.touringplans.com/datasets/flight_of_passage.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
dinosaur.hour <- cleanthis("https://cdn.touringplans.com/datasets/dinosaur.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
ee.hour <- cleanthis("https://cdn.touringplans.com/datasets/expedition_everest.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
safari.hour <- cleanthis("https://cdn.touringplans.com/datasets/kilimanjaro_safaris.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
navi.hour <- cleanthis("https://cdn.touringplans.com/datasets/navi_river.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
pirate.hour <- cleanthis("https://cdn.touringplans.com/datasets/pirates_of_caribbean.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
rnrc.hour <- cleanthis("https://cdn.touringplans.com/datasets/rock_n_rollercoaster.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
sevendwarf.hour <- cleanthis("https://cdn.touringplans.com/datasets/7_dwarfs_train.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
slinkydog.hour <- cleanthis("https://cdn.touringplans.com/datasets/slinky_dog.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
spaceship.hour <- cleanthis("https://cdn.touringplans.com/datasets/spaceship_earth.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
splash.hour <- cleanthis("https://cdn.touringplans.com/datasets/splash_mountain.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
toystory.hour <- cleanthis("https://cdn.touringplans.com/datasets/toy_story_mania.csv", by="hour") %>%
  mutate(month = month(datetime), year = year(datetime) )
 
whichyear <- 2021
soarin.hour <- soarin.hour %>% filter(year==whichyear) %>% select( SPOSTMIN)%>% rename(soarin = SPOSTMIN)%>%
  pivot_longer(.,cols= "soarin",names_to = "Ride", values_to = "value")

avatar.hour <- avatar.hour %>% filter(year==whichyear) %>% select( SPOSTMIN)%>% rename(avatar = SPOSTMIN)%>%
  pivot_longer(.,cols= "avatar",names_to = "Ride", values_to = "value")
  
dinosaur.hour <- dinosaur.hour %>% filter(year==whichyear) %>% select( SPOSTMIN)%>% rename(dinosaur = SPOSTMIN)%>%
  pivot_longer(.,cols= "dinosaur",names_to = "Ride", values_to = "value")
  
ee.hour <- ee.hour %>% filter(year==whichyear) %>% select( SPOSTMIN)%>% rename(ee = SPOSTMIN)%>%
  pivot_longer(.,cols= "ee",names_to = "Ride", values_to = "value")

safari.hour <- safari.hour %>% filter(year==whichyear) %>% select( SPOSTMIN)%>% rename(ee = SPOSTMIN)%>%
  pivot_longer(.,cols= "ee",names_to = "Ride", values_to = "value")

navi.hour <- navi.hour %>% filter(year==whichyear) %>% select( SPOSTMIN)%>% rename(navi = SPOSTMIN)%>%
  pivot_longer(.,cols= "navi",names_to = "Ride", values_to = "value")

pirate.hour <- pirate.hour %>% filter(year==whichyear) %>% select( SPOSTMIN)%>% rename(pirate = SPOSTMIN)%>%
  pivot_longer(.,cols= "pirate",names_to = "Ride", values_to = "value")

rnrc.hour <- rnrc.hour %>% filter(year==whichyear) %>% select( SPOSTMIN)%>% rename(rnrc = SPOSTMIN)%>%
  pivot_longer(.,cols= "rnrc",names_to = "Ride", values_to = "value")

sevendwarf.hour <- sevendwarf.hour %>% filter(year==whichyear) %>% select( SPOSTMIN)%>% rename(sevendwarf = SPOSTMIN)%>%
  pivot_longer(.,cols= "sevendwarf",names_to = "Ride", values_to = "value")

slinkydog.hour <- slinkydog.hour %>% filter(year==whichyear) %>% select( SPOSTMIN)%>% rename(slinkydog = SPOSTMIN)%>%
  pivot_longer(.,cols= "slinkydog",names_to = "Ride", values_to = "value")

spaceship.hour <- spaceship.hour %>% filter(year==whichyear) %>% select( SPOSTMIN)%>% rename(spaceship = SPOSTMIN)%>%
  pivot_longer(.,cols= "spaceship",names_to = "Ride", values_to = "value")

splash.hour <- splash.hour %>% filter(year==whichyear) %>% select( SPOSTMIN)%>% rename(splash = SPOSTMIN)%>%
  pivot_longer(.,cols= "splash",names_to = "Ride", values_to = "value")

toystory.hour<- toystory.hour %>% filter(year==whichyear) %>% select( SPOSTMIN)%>% rename(toystory = SPOSTMIN)%>%
  pivot_longer(.,cols= "toystory",names_to = "Ride", values_to = "value")

allinone <- rbind(soarin.hour, avatar.hour,dinosaur.hour, ee.hour, safari.hour ,navi.hour, pirate.hour, rnrc.hour, sevendwarf.hour, slinkydog.hour, spaceship.hour, splash.hour, toystory.hour) 
mul.box2 <- allinone %>% 
  ggplot(aes(x=reorder(Ride,-value, median), y=value))+
  geom_boxplot(varwidth = TRUE)+
  ggtitle("Multiple boxplot of all rides 2021")+
  xlab("Ride name")+
  ylab("Waiting Time")+
  theme_grey(14)+
  theme(axis.text.x=element_text(size=9))

=======
>>>>>>> 6368b076be23142e42c2efbd6b84ce91ab284f7a
```


```{r, include=FALSE}
#Create empty time data frame to match the rows of all attractions because the data files for each attraction may have NA values for some days
timedf <- data.frame(date = seq(as.Date('2015-01-01'), as.Date('2021-12-28'), by = 'days'))
weekofmonth <- function(date){
  out <- as.integer(strftime(date, format = "%U"))-as.integer(strftime(floor_date(date, unit="months"), format = "%U"))+1
  return(out)
}
timedf$year <- year(timedf$date)
timedf$month <- month(timedf$date)
timedf$weekofmonth <- weekofmonth(timedf$date)
timedf$weekday <- wday(timedf$date,label = TRUE)
timedf$weekend <- ifelse(timedf$weekday %in% c("Sat","Sun") , TRUE, FALSE)
```

```{r}
######Overall dataset of Daily time of all attractions 
######Use this one!
attractions <- left_join(timedf, soarin) %>%left_join(., avatar) %>% 
  left_join(., dinosaur) %>% left_join(., ee) %>% 
  left_join(., safari) %>% left_join(., navi) %>% 
  left_join(., pirate) %>% left_join(., rnrc) %>% 
  left_join(., sevendwarf) %>% left_join(., slinkydog) %>% 
  left_join(., spaceship) %>% left_join(., splash) %>% 
   left_join(., toystory)
head(attractions)
```

## Entities dataset

For the entities dataset, we add a new column `ratio` to the dataset by calculating waiting time/duration.

Besides, the orignal name of the attractions is very long, so we use the `short name` column instead of the `name` column. After that, we filter out the column that is useful for the visualization (short_name, land, ratio).

```{r}
#load entity dataset
entities <- read_csv("https://cdn.touringplans.com/datasets/entities.csv")
ratio_data<-entities %>% 
  mutate(`ratio`=round(average_wait_per_hundred/duration,2)) %>% 
  select(short_name,land,ratio) %>% 
  drop_na() %>% 
  distinct(short_name,.keep_all = TRUE)
head(ratio_data)
```


```{r, include=FALSE}
#Soarin, 2021 Apr data, by weekend/weekday
soarin.hour <- cleanthis("https://cdn.touringplans.com/datasets/soarin.csv", by="hour")

soarin.hour$year <- factor(year(soarin.hour$date))
soarin.hour$month <- factor(month(soarin.hour$date))
#timedf$weekofmonth <- weekofmonth(timedf$date)
soarin.hour$weekday <- factor(wday(soarin.hour$date,label = TRUE))
soarin.hour$weekend <- ifelse(soarin.hour$weekday %in% c("Sat","Sun") , TRUE, FALSE)
soarin.hour <- left_join(soarin.hour, bigmeta %>% select(DATE, HOLIDAY) %>% rename(date=DATE))
soarin.hour %>% filter(year==2021 & month==4) %>% group_by(date,weekday) %>%
  summarize(mean=mean(SPOSTMIN)) 
soarin_byhour <- soarin.hour  %>% filter(year==2021) %>% group_by(weekend,hour) %>% summarize(mean = mean(SPOSTMIN))

soarin_dt <- data.frame(hour = as.character(seq(10,23)))

soarin_dt <- left_join(soarin_dt, soarin_byhour %>% filter(weekend==FALSE) %>% select(hour, mean) %>%  rename(weekday = mean) %>% ungroup(), by="hour") %>% select(-weekend)
soarin_dt <- soarin_dt %>% left_join(., soarin_byhour %>% filter(weekend==TRUE) %>%rename(Weekend_mean = mean), by="hour") %>% select(-weekend)
<<<<<<< HEAD
soarin_dt

barplot(soarin_byhour$mean)
=======
>>>>>>> 6368b076be23142e42c2efbd6b84ce91ab284f7a

write.csv(soarin_dt, file="soarin_int.csv")
```

<<<<<<< HEAD
```{r}
#attractions split by weekend and month, analyze whether weekends have longer or shorter waiting time for all rides
filtered <- attractions %>% filter(year>=2019) %>% group_by(weekend) %>% summarise(across(c(soarin:pirate),mean, na.rm=TRUE))

ggplot(filtered)
test <- filtered %>% pivot_longer(cols=!"weekend", values_to = "time", names_to = "ride") %>%filter(weekend==FALSE) %>% factor(ride)
test$ride <- factor(test$ride)
  
```

```{r}
=======
```{r, include=FALSE}
>>>>>>> 6368b076be23142e42c2efbd6b84ce91ab284f7a
#Suppose we want to use Jul 2021 data to see what the waiting time will be if we go this month
#We selected three classic rides in Magic Kingdom: Pirates of the Caribbean, Seven Dwarfs Mine Train, Splash Mountain

#By hour data, if needed
pi_hour <- cleanthis("https://cdn.touringplans.com/datasets/pirates_of_caribbean.csv", by="hour") 

sd_hour <- cleanthis("https://cdn.touringplans.com/datasets/7_dwarfs_train.csv", by="hour")

splash_hour <- cleanthis("https://cdn.touringplans.com/datasets/splash_mountain.csv", by="hour")

ride3 <- attractions %>% select(date, year, month, weekday, weekend, pirate, sevendwarf, splash) %>% filter(year==2021 & month ==7) 
ride3 <- ride3 %>% left_join(., bigmeta %>% filter(YEAR==2021 & MONTHOFYEAR==7) %>% select(DATE, WDWMEANTEMP) %>% rename(date=DATE))

ride3 %>% summarize(mean1=mean(pirate), mean2=mean(sevendwarf), mean3=mean(splash))
```

```{r, include=FALSE}
##daily information of selected attractions

ride3
try <- pi_hour %>% mutate(year = year(date), month = month(date), weekday = wday(date, label=TRUE), weekend=ifelse(weekday %in% c("Sat","Sun") , TRUE, FALSE)) %>% filter(year==2021 & month==7) %>% group_by(date,hour) %>% summarize(mean=mean(SPOSTMIN))

try <- try %>% pivot_wider( id_cols="date",id_expand=FALSE, values_from = "mean", names_from = "hour")
```



  
```{r, include=FALSE}
# data transformation for barchart (for events)
events_data <- bigmeta %>% 
  select(DATE,WDWeventN,MKeventN,EPeventN,HSeventN,AKeventN) %>% 
  pivot_longer(!DATE,names_to = "Park_name", values_to = "events") %>% 
  drop_na("events") %>% 
  mutate("Park_name"=replace(Park_name,Park_name=="WDWeventN","Walt Disney World"),
         "Park_name"=replace(Park_name,Park_name=="MKeventN","Magic Kingdom"),
         "Park_name"=replace(Park_name,Park_name=="EPeventN","Epcot"),
         "Park_name"=replace(Park_name,Park_name=="HSeventN","Hollywood Studios"),
         "Park_name"=replace(Park_name,Park_name=="AKeventN","Animal Kingdom")
  )
```


```{r, include=FALSE}
# data transformation for time series line plot
attractions_ts<-attractions %>% 
  filter(year=="2019") %>% 
  select(-c(year,month, weekday,weekend)) %>% 
  pivot_longer(-date,names_to = "attraction", values_to = "waiting_time") %>% 
  mutate(attraction=fct_reorder(attraction,waiting_time,median),attraction=fct_relevel(attraction,"ee",after=4))
```


```{r, include=FALSE}
# data transformation for mosaic plot
mosaic_ticketprice<-bigmeta %>% 
  drop_na(DAYOFWEEK,WDW_TICKET_SEASON) %>% 
  group_by(DAYOFWEEK,WDW_TICKET_SEASON) %>% 
  rename(ticket_price=WDW_TICKET_SEASON) %>% 
  summarise(Freq=n())
```
